Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-20 paper/paper.tex
update
in the middle of revising the Probabilistic Program section

Anglican is a language for expressing probabilistic programs.
A probabilistic program is a stateful \textit{deterministic} computation
$\mathcal{P}$ with the following properties~\cite{TMP+15}:
\begin{itemize}
\item Initially, $\mathcal{P}$ expects no arguments.
\item On every call, $\mathcal{P}$ returns either a distribution
  and an address $(F,\alpha)$, a distribution and a value $(G,
        y)$, or a value $z$.
\item Upon returning $F$, $\mathcal{P}$ expects a value $x$ drawn from $F$ as
    the argument to the next call (\texttt{sample} in Anglican).
\item Upon returning $(G, y)$, $\mathcal{P}$ is invoked again
    without arguments (\texttt{observe} in Anglican).
\item Upon returning $z$, $\mathcal{P}$ terminates.
\end{itemize}
A program is run by calling $\mathcal{P}$ repeatedly until
termination.

A program need not generate the same sequence of random variables in every
execution. For this reason we assume that each random variable $x$ is assigned
a unique label $\alpha$, which we call an address, that induces a
correspondence between variables in different executions. Every execution
implicitly produces a sequence of triples of distributions, values of
\textit{latent} random variables, and addresses, $(F_i, x_i, \alpha_i)$.  We
call this sequence a \textit{trace} and denote it by $\pmb{x}$. A trace induces
a sequence of pairs $(G_j, y_j)$ of distributions and values of
\textit{observed} random variables. We call this sequence an \textit{image} and
denote it by $\pmb{y}$. For notational simplicity we assume that the program
always generates the same ordered set $\pmb{z}$ of \textit{results} $z_k$.

@@

>>>>

\hy{Need to be revised.}

Programs in Anglican define probability distributions 
over sequences of sampled values through their executions.

Imagine an interpreter of Anglican that runs the deterministic parts of a Anglican program 
according to its standard semantics, executes its \texttt{sample} form by
generating a random sample, and treats its $\texttt{observe}$ form by skip.
In addition to this execution, the interpreter keeps a log that records
information about all the \texttt{sample} and \texttt{observe} forms encountered 
during execution. The information kept for \texttt{sample} is a triple $(F,x,\alpha)$ of
the following data:
\begin{itemize}
        \item $F$ is one of the built-in primitive probability distributions, 
                such as the standard normal, for which we have
                the probability density or probability mass function $p_F$.
        \item $x$ is a value sampled from the distribution $F$.
        \item $\alpha$ is an address that uniquely and systematically identifies the random 
                choice made; this address is exploited in inference algorithms, as
                explained later.
\end{itemize}
The information kept for \texttt{observe} is a pair $(G,y)$ where $G$
is a primitive probability distribution as $F$ above and $y$ is an observed value.
Thus, a log is just a sequence of triples $(F,x,\alpha)$ and pairs $(G,y)$.

We call the triple part of a log \textit{trace}. This part records all the random 
choices of an execution. Thus, if the trace of a log is fixed, the rest of the log 
is decided determistically by the program. A trace induces
a sequence of pairs $(G_j, y_j)$ of distributions and values of
\textit{observed} random variables. We call this sequence an \textit{image} and
denote it by $\pmb{y}$. For notational simplicity we assume that the program
always generates the same ordered set $\pmb{z}$ of \textit{results} $z_k$.

@@

%\section{Probabilistic Program}
%\label{sec:pp-def}
%
%Anglican is a language for expressing probabilistic programs.
%A probabilistic program is a stateful \textit{deterministic} computation
%$\mathcal{P}$ with the following properties~\cite{TMP+15}:
%\begin{itemize}
%\item Initially, $\mathcal{P}$ expects no arguments.
%\item On every call, $\mathcal{P}$ returns either a distribution
%  and an address $(F,\alpha)$, a distribution and a value $(G,
%        y)$, or a value $z$.
%\item Upon returning $F$, $\mathcal{P}$ expects a value $x$ drawn from $F$ as
%    the argument to the next call (\texttt{sample} in Anglican).
%\item Upon returning $(G, y)$, $\mathcal{P}$ is invoked again
%    without arguments (\texttt{observe} in Anglican).
%\item Upon returning $z$, $\mathcal{P}$ terminates.
%\end{itemize}
%A program is run by calling $\mathcal{P}$ repeatedly until
%termination.
%
%A program need not generate the same sequence of random variables in every
%execution. For this reason we assume that each random variable $x$ is assigned
%a unique label $\alpha$, which we call an address, that induces a
%correspondence between variables in different executions. Every execution
%implicitly produces a sequence of triples of distributions, values of
%\textit{latent} random variables, and addresses, $(F_i, x_i, \alpha_i)$.  We
%call this sequence a \textit{trace} and denote it by $\pmb{x}$. A trace induces
%a sequence of pairs $(G_j, y_j)$ of distributions and values of
%\textit{observed} random variables. We call this sequence an \textit{image} and
%denote it by $\pmb{y}$. For notational simplicity we assume that the program
%always generates the same ordered set $\pmb{z}$ of \textit{results} $z_k$.
%
%The target density $\pi(\vec{x}) := \gamma(\vec{x}) / Z$ of a program
%is defined in terms of the product of the probabilities of all random
%choices $\pmb{x}$ and the likelihood of all observations $\pmb{y}$
%\begin{align}
%  \gamma(\pmb{x}) 
%  &:= 
%%  \prod_{i=1}^{\left|\pmb{x}\right|}
%  p_{F_i}(x_i) \prod_{j=1}^{\left|\pmb{y}\right|}p_{G_j}(y_{j}).
%  \label{eqn:p-trace}
%\end{align}
%The objective of inference in probabilistic program $\mathcal{P}$
%is to discover the distribution of $\pmb{z}$.
